The Shape Matching Element Method: Direct Animation of Curved Surface Models

### 2021-01-27 paper/sections/methods.tex
section 5 first pass done

We begin by describing shape matching to a single NURBS part. 
Given a two different configurations of the same part (specified by the control points), shape matching computes a polynomial deformation which best aligns them.

Let $\vc{q^0}$ be the initial control points of the part, provided by the modeller and Let $\vc{q}$ be a vector of modified control point values.
We can compute the undeformed (or reference) position of any point on our part using $\refX\left(u,v\right) = \Jnurbs\left(u,v\right)\mathbf{q^0}$ and the corresponding
deformed (world space) position as $\defX\left(u,v\right) = \Jnurbs\left(u,v\right)\mathbf{q}$.

@@

\dave{Ty: need some text to explain how we pick points on the nurbs surface.}
\ty{Need to think a little more on how to write this. The main idea is that each surface has a knot vector and each Bspline covers a span of those knots. Therefore to guarantee we represent all control points for a surface, we uniformly sample $d+1$ points in each span, where $d$ is the NURBS surface}
We discretize the shape matching energy and minimize by solving
\begin{equation}
\samp{\Pmat}^T\wt\samp{\Pmat}\Pcox = \samp{\Pmat}^T\wt\samp{\Jnurbs}\vc{q}.
\label{eq:shapematch}
\end{equation} Here $\samp{\Pmat}$ (resp. $\samp{\Jnurbs}$) is a $3s \times 3k$ (resp. $3s \times 3nm$) matrix where $s$ is the number of quadrature points, 
$k$ is the number of  terms in the shape matching polynomial and $3nm$ is the number of NURBS control points. 
$\samp{\Pmat}$ (resp. $\samp{\Jnurbs}$) is constructed by stacking the evaluations of $\Pmatxuv$ (resp. $\Jnurbs\left(u, v\right)$) at each
quadrature point.
$\wt$ is a diagonal, $3s \times 3s$ matrix of integration weights.

\paragraph*{Relationship to the Virtual Element Method}
For a non-planar NURBS part, and given a sufficient number of quadrature points, ~\refeq{shapematch} will be well-posed allowing us to write: $\Pcox^* = \Pi\vc{q}$ where
\begin{equation*}
    \Pi = \left(\samp{\Pmat}^T\wt\samp{\Pmat}\right)^{-1}\samp{\Pmat}^T\wt\samp{\Jnurbs},
\end{equation*} is a projection from the space of control points onto the space of polynomials. 
This projection serves an identical purpose (and is mathematically equivalent up to choice of metric) to the projection operator used in VEM~\cite{10.1142/S021820251440003X}.
VEM uses a single polynomial projection per polyhedral element to solve differential equations on complex volumetric meshes.
We interpret shape matching as converting our NURBS part into a single Virtual Element suitable for elastodynamics simulation.
Unfortunately a single polynomial deformation space will not be suitably expressive for even moderately complex models. 
Previous work~\cite{10.1145/1073204.1073216,10.1145/1275808.1276480} used additional lattices or partitions of the reference space to 
increase the number of polynomials in the shape matching. Next we outline our meshless Shape Element Method which works directly on the NURBS
geometry itself.

@@

 \begin{equation}
    \defX\left(\refX\right) = \sum_{\prt_i\in\Prts} \bld_i\Pmatxi\Pi_i\vc{q}_i, 
    \label{eq:kinematics}
 \end{equation} where $w_i$ $\Pmatx_i$, $\Pi_i$ and $\vc{q}_i$ are the blending weights, polynomial basis, projection operator and control points for the $i^{th}$
 part respectively.

>>>>

We begin by describing shape matching to a single NURBS part. 
Given two different configurations of the same part (specified by the control points), shape matching computes a polynomial deformation that best aligns them.

Let $\vc{q^0}$ be the initial control points of the part, provided by the modeller, and $\vc{q}$ be a vector of modified control point values.
We can compute the undeformed (or reference) position of any point on our part using $\refX\left(u,v\right) = \Jnurbs\left(u,v\right)\mathbf{q^0}$ and the corresponding
deformed (world space) position as $\defX\left(u,v\right) = \Jnurbs\left(u,v\right)\mathbf{q}$.

@@

We discretize the shape matching energy and minimize by solving
\begin{equation}
\samp{\Pmat}^T\wt\samp{\Pmat}\Pcox = \samp{\Pmat}^T\wt\samp{\Jnurbs}\vc{q}.
\label{eq:shapematch}
\end{equation} Here $\samp{\Pmat}$ (resp. $\samp{\Jnurbs}$) is a $3s \times 3(k+1)$ (resp. $3s \times 3nm$) matrix where $s$ is the number of quadrature points, 
$k$ is the number of  terms in the shape matching polynomial and $3nm$ is the number of NURBS control points. 
$\samp{\Pmat}$ (resp. $\samp{\Jnurbs}$) is constructed by stacking the evaluations of $\Pmatxuv$ (resp. $\Jnurbs\left(u, v\right)$) at each
quadrature point.
$\wt$ is a diagonal, $3s \times 3s$ matrix of integration weights.

Quadrature points are selected so that $\samp{\Jnurbs}$ is of sufficient rank required to represent all control points. We achieve this by sampling over each knot span defined by the  B-splines, and find it effective to use the order of the NURBS as the number of quadrature points sampled in each knot span.
\ty{the number of samples may be different if order in the u direction and v differ, but I figured it was best to keep this simple.}

\paragraph*{Relationship to the Virtual Element Method}
For a non-planar NURBS part, and given a sufficient number of quadrature points, ~\refeq{shapematch} will be well-posed allowing us to write: $\Pcox^* = \Pi\vc{q}$ where
\begin{equation*}
    \Pi = \left(\samp{\Pmat}^T\wt\samp{\Pmat}\right)^{-1}\samp{\Pmat}^T\wt\samp{\Jnurbs},
\end{equation*} is a projection from the space of control points onto the space of polynomials. 
This projection serves an identical purpose (and is mathematically equivalent up to choice of metric) to the projection operator used in VEM~\cite{10.1142/S021820251440003X}.
VEM uses a single polynomial projection per polyhedral element to solve differential equations on complex volumetric meshes.
We interpret shape matching as converting our NURBS part into a single Virtual Element suitable for elastodynamics simulation.
Unfortunately, a single polynomial deformation space will not be suitably expressive for even moderately complex models.
Previous work~\cite{10.1145/1073204.1073216,10.1145/1275808.1276480} used additional lattices or partitions of the reference space to 
increase the number of polynomials in the shape matching. Next we outline our meshless Shape Matching Element Method which works directly on the NURBS
geometry itself.

@@

 \begin{equation}
    \defX\left(\refX\right) = \sum_{\prt_i\in\Prts} \bldxi\Pmatxi\Pi_i\vc{q}_i, 
    \label{eq:kinematics}
 \end{equation} where $\bldxi$, $\Pmatx_i$, $\Pi_i$ and $\vc{q}_i$ are the blending weights, polynomial basis, projection operator and control points for the $i^{th}$
 part respectively.
